
#Prediction Assignment Writeup

```{r setup, echo=FALSE}
# set global chunk options: 
library(knitr)
opts_chunk$set(cache=TRUE, autodep = TRUE)
dep_auto()
```
 

```{r}
require(caret)
require(ElemStatLearn)
require(pgmm)
require(rpart)
require(gbm)
require(lubridate)
require(forecast)
require(e1071)
require(rattle)
require(party)
```

```{r}
#Reading the training dataset
data = read.csv("C:/Users/Isadora/Desktop/pml-training.csv")
dim(data)
str(data)
```

```{r}
#Partitioning the training dataset into training, testing and validation

splitSample <- sample(1:3, size=nrow(data), prob=c(0.7,0.15,0.15), replace = TRUE)
training <- data[splitSample==1,]
testing <- data[splitSample==2,]
validation <- data[splitSample==3,]
dim(training)
dim(testing)
dim(validation)
```

```{r}
#Checking distribuition of the covariates
summary(training)
```

```{r}
#Selecting only covariates with no nulls and covariates that do not
#represent some aggregated function (like kurtosis, amplitude, etc)

training <- training[,c("roll_belt","pitch_belt","yaw_belt","total_accel_belt","gyros_belt_x",
"gyros_belt_y","gyros_belt_z","accel_belt_x","accel_belt_y","accel_belt_z",
"magnet_belt_x","magnet_belt_y","magnet_belt_z","roll_arm","pitch_arm","yaw_arm",
"total_accel_arm","gyros_arm_x","gyros_arm_y","gyros_arm_z","accel_arm_x",
"accel_arm_y","accel_arm_z","magnet_arm_x","magnet_arm_y","magnet_arm_z",
"roll_dumbbell","pitch_dumbbell","yaw_dumbbell","total_accel_dumbbell",
"gyros_dumbbell_x","gyros_dumbbell_y","gyros_dumbbell_z","accel_dumbbell_x",
"accel_dumbbell_y","accel_dumbbell_z","magnet_dumbbell_x","magnet_dumbbell_y",
"magnet_dumbbell_z","roll_forearm","pitch_forearm","yaw_forearm","total_accel_forearm",
"gyros_forearm_x","gyros_forearm_y","gyros_forearm_z","accel_forearm_x",
"accel_forearm_y","accel_forearm_z","magnet_forearm_x","magnet_forearm_y",
"magnet_forearm_z","classe")]
```
##Let`s take a look into the covariates
```r
featurePlot(x=training, y=training$classe, plot="pairs")
```

```{r}
qplot(roll_belt, classe, data=training)
```

##Fitting CART
```{r}
set.seed(567)
fit_cart <- train(classe ~., data = training, method = "rpart")
print(fit_cart$finalModel)
fancyRpartPlot(fit_cart$finalModel)
#Accuracy in training
confusionMatrix(training$classe,predict(fit_cart,training))$overall[1]
#Resulting accuracy = 0.4978
#Accuracy in testing
confusionMatrix(testing$classe,predict(fit_cart,testing))$overall[1]
#Resulting accuracy = 0.4952
```

##Fitting Ctree
```{r}
set.seed(890)
fit_ctree <- train(classe ~., data = training, method = "ctree2")
print(fit_ctree)
print(fit_ctree$finalModel)
plot(fit_ctree$finalModel, type = "simple")
#Accuracy in training
confusionMatrix(training$classe,predict(fit_ctree,training))$overall[1]
#Resulting accuracy = 0.3922
#Accuracy in testing
confusionMatrix(testing$classe,predict(fit_ctree,testing))$overall[1]
#Resulting accuracy = 0.3767
```

##Fitting boosted tree
```r
#Cross validation will be performed here in order to avoid overfitting
set.seed(765)
traincontrol <- trainControl(method="cv", number=10)
fit_boost <- train(classe ~., data = training, method = "gbm",
			 trControl = traincontrol, verbose=FALSE)
#Could not be performed because R crashed. My pc is too slow and poor
```

##Fitting random forest
```r
#Cross validation will be performed here in order to avoid overfitting
set.seed(234)
traincontrol <- trainControl(method="cv", number=10)
fit_rf <- train(classe ~., data = training, method = "rf", 
		    trControl = traincontrol, prox=TRUE)
#Could not be performed because R crashed. My pc is too slow and poor
```

##Fitting random forest with most important predictors (above 50%) of fitted CART and Ctree
```{r}
varImp(fit_cart)
varImp(fit_ctree)
```

```r
#Cross validation will be performed here in order to avoid overfitting
set.seed(234)
traincontrol <- trainControl(method="cv", number=10)
fit_rf <- train(classe ~ pitch_forearm + roll_belt + roll_forearm + magnet_dumbbell_y + accel_forearm_x + magnet_arm_x, 
			data = training, method = "rf", 
		    	trControl = traincontrol, ntree = 3, prox=TRUE)
#Could not be performed because R crashed. My pc is too slow and poor
```

##Combining predictors of CART and Ctree (which was the only models that could be performed!)
```{r}
#Predictions in testing set
pred_cart <- predict(fit_cart, newdata=testing)
pred_ctree <- predict(fit_ctree, newdata=testing)
#Combining predictors
df <- data.frame(pred_cart,pred_ctree,classe=testing$classe)
set.seed(579)
fit_combined <- train(classe ~., data = df, method = "rf")
pred_combined <- predict(fit_combined, df)
#Accuracy in testing
confusionMatrix(testing$classe,predict(fit_combined,testing))$overall[1]
#Accuracy = 0.5404
```

##Checking the accuracy of this final model on the validation set
```{r}
#Accuracy in validation
#Remember fit_cart and fit_ctree
pred_cart_validation <- predict(fit_cart, validation)
pred_ctree_validation <- predict(fit_ctree, validation)
validation2 <- data.frame(pred_cart_validation,pred_ctree_validation,classe=validation$classe)
pred_combined_validation <- predict(fit_combined, validation2)
confusionMatrix(validation2$classe,predict(fit_combined,validation2))$overall[1]
#Accuracy = 0.5304
```



